---
output: github_document
---

<!-- README.md is generated from the README.Rmd. Please edit that file -->

# RSV concentrations in Central California, USA
This is the code to replicate the results in Section 3.2 of the manuscript. You will need the following libraries. The R version used for this analysis and the package versions are listed below. 

```{r setup, warning=FALSE, message=FALSE}
# > sessionInfo()
# R version 4.4.1 (2024-06-14) -- "Race for Your Life"
# Platform: aarch64-apple-darwin20
# Running under: macOS Monterey 12.4

# To install OSplines package
# install.packages("remotes")
# remotes::install_github("AgueroZZ/OSplines")

library(OSplines) # OSplines_0.1.1
library(aghq) # aghq_0.4.1
library(readr) # readr_2.1.5
library(dplyr) # dplyr_1.1.4
library(magrittr) # magrittr_2.0.3 
library(lubridate) # lubridate_1.9.3 
library(Matrix) # Matrix_1.7-0 
library(TMB) # TMB_1.9.14 
library(ggplot2) # ggplot2_3.5.1
library(janitor) # janitor_2.2.0
```

## Data
The RSV concentration data is found in the data folder and it was download from https://purl.stanford.edu/cx529np1130. Set your working directory to Section 3.2 and run the following code:

```{r, warning=FALSE, message=FALSE}
raw_d <- read_csv("./data/SCAN_AllPlants_SDR_7April23_rev.csv")

# creating a sample_date and renaming the station column
raw_d <- raw_d %>% 
  mutate(sample_date = ymd(paste0(Year,"-",Month,"-",Day))) %>% 
  rename(site_id = `Plant Abbr`) 

# clean column names (might not need it's my habit)
work_d <- raw_d %>% clean_names() 
```

Here is a plot of the raw data, note if a “0” appears, it means the assay was a non-detect. The detection limit varies by sample depending on the amount of solids by dry weight included, but is between 500–1000 cp/g dry weight. RSV measurements begin on January 5th, 2022. 

```{r, warning=FALSE, out.width="75%"}
ggplot(raw_d %>% filter(sample_date >= "2022-01-05"), aes(sample_date,RSV_gc_g_dry_weight))+ 
  geom_point(size=0.5)+
  facet_wrap(~site_id, nrow = 3)+ 
  scale_x_date(breaks=scales::pretty_breaks(n=5), name = "",date_labels ="%b")+
  theme_bw()
```

## Prepping Data for Modeling
The function "prep_data" will prepare the data for modeling. Here is an example of how to use this function is found below. 

- Important Notes about Normalization: 
  - If no normalization variable is used, then create a denominator column with values of ones 
    - Example: "mutate(denom = 1)" and denom_column_name = "denom" in prep_data( ) function
  - In this example, we will use PMMoV t normalize RSV concentrations. 
  - Both PMMoV and RSV concentrations are measured as gc/g and therefore the PMMoV-normalized RSV concentrations are unitless. 
  
- Important Notes about Censoring: 
  - If no outcome values are censored, create a censored_y column with values of FALSE
    - Example: mutate(censored_y = FALSE)
  - If outcome values are censored, do the two following things: 
    - Set the censored outcome values to the detection limit. 
    - create a censored_y column with values of FALSE if the outcome is observed and values of TRUE if the outcome is censored. 
  - In this example, we have censored values and we will use the detection limit of 1000 cp/g.

```{r, eval = FALSE}

# In this example, 
## pm_mo_v_gc_g_dry_weight is used for normalization
## the outcome is RSV_gc_g_dry_weight
## RSV_gc_g_dry_weight values are set to 0 if censored

# So following the censored instructions above: 
work_d < - work_d %>% 
  mutate(censored_y = ifelse(RSV_gc_g_dry_weight == 0, TRUE, FALSE), # create censored_y column
         RSV_gc_g_dry_weight = ifelse(y==0, 1000, y))                # set censored values to limit

tmbdat <- prep_date(outcome_column_name = "",
                    denom_column_name = "pm_mo_v_gc_g_dry_weight")  

```

## Prior Specifications


